---
title: 'Soil Eosion in California County'
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
---


```{r setup, include=FALSE, echo=FALSE}

knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE,
    error = FALSE, 
    dpi = 400
)

```


# Setup

## Load libraries

```{r}
library(tidyverse)
library(sf)
library(raster)
library(usdata)
library(tigris)
library(skimr)
library(DataExplorer)
library(raster)
library(terra)
```


# Soil Erosion

RUSLE (A = R x K x LS x C x P)

## Resolutions

- Rainfall = 4 km
- Erodibility = 30 m
- Slope = 10 m
- Crop scape = 30


## Rainfall Erosivity

### Load data

```{r, eval=FALSE, include=TRUE}
# rainfall - 2021
rnf <- stack("Data/Rainfall/dataset-sis-biodiversity-cmip5-global-83e8ca0e-4eea-4051-a911-c3299d0f2380 (1)/BIO12_gfdl-esm2m_rcp85_r1i1p1_1950-2100_v1.0.nc")
rnf <- mean(rnf[[72]])*3600*24*365*1000
plot(rnf)
```


### Plot


```{r, eval=FALSE, include=TRUE}
#plot(rnf)
#sum_rnf <- sum(rnf)

plot(rnf)
plot(st_geometry(california), col = "red", add = TRUE)
```


### Cropping and masking

#### Boundary lines

```{r}
library(tigris)

# california
california <- counties(cb = TRUE, progress_bar = FALSE) %>% 
  filter(STATE_NAME == "California")

sj <- counties(cb = TRUE, progress_bar = FALSE) %>% 
  filter(NAME == "San Joaquin")

plot(st_geometry(california))
plot(st_geometry(sj), col = "blue", add = TRUE)
```


```{r, eval=FALSE, include=TRUE}
# cropping
crop_rnf <- raster::crop(rnf, extent(california))
plot(crop_rnf)

# masking
mask_rnf <- raster::mask(crop_rnf, mask = california)
plot(mask_rnf)

# resampling/interpolation
target <- mask_rnf
res(target) <- 0.0003176353
res_rnf <- raster::resample(mask_rnf, target, method = "bilinear")
plot(res_rnf)

raster::writeRaster(res_rnf, filename = file.path("Data/Rainfall/PRISM//res_rnf_california.tif"), format="GTiff", overwrite=TRUE)
```


```{r, eval=FALSE, include=TRUE}
res_rnf <- raster("Data/Rainfall/PRISM/res_rnf_california.tif")
plot(res_rnf)
plot(st_geometry(sj), add = TRUE)

# cropping
crop_rnf <- raster::crop(res_rnf, extent(sj))
plot(crop_rnf)

# masking
mask_rnf <- raster::mask(crop_rnf, mask = sj)
plot(mask_rnf)

raster::writeRaster(mask_rnf, filename = file.path("Data/Rainfall/PRISM//res_rnf_sj.tif"), format="GTiff", overwrite=TRUE)
```

### Erosivity factor

```{r}
# load data 
rnf_sj <- raster("Data/Rainfall/res_rnf_sj.tif")
plot(rnf_sj)

# The equation of Moore
# Implement the equation of Moore as a function in R
calculate_r_moore <- function(p) {
  ke <- 11.46*p - 2226
  r <- 0.029*ke - 26
  r_si <- 17.02*r # Conversion from imperial to SI units
  return(r_si)
}

r_moore <- calc(x = rnf_sj, fun = calculate_r_moore)
plot(r_moore)

# save raster
raster::writeRaster(r_moore, filename = file.path("Data/Rainfall/erosivity_factor.tif"), format="GTiff", overwrite=TRUE)
```



## Erodibility factor (K)

### Load data

```{r}
ef <- raster("Data/Erodibility/SJ_Erodibility.tif")
raster::plot(ef)
```


## Slope length and steepness


### DEM

#### Load data

```{r, eval=FALSE, include=TRUE}
dem1 <- raster("Data/DEM or DSM/USGS_13_n38w121_20220103.tif")
dem2 <- raster("Data/DEM or DSM/USGS_13_n38w122_20220810.tif")
dem3 <- raster("Data/DEM or DSM/USGS_13_n39w121_20220206.tif")
dem4 <- raster("Data/DEM or DSM/USGS_13_n39w122_20220206.tif")

dem_2022 <- raster::mosaic(dem1, dem2, dem3, dem4,  fun = 'mean')
plot(dem_2022)
#plot(st_geometry(california), add = TRUE)
plot(st_geometry(sj), add = TRUE, col = "red")


sj =sf::st_set_crs(sj, crs(dem_2022))
plot(dem)
plot(st_geometry(sj), add = TRUE)
```


### Mosaic

```{r, eval=FALSE, include=TRUE}
dem_2022 <- raster::mosaic(dem1, dem2, dem3, dem4,  fun = 'mean')

# plot DEM
plot(dem_2022)
plot(st_geometry(sj), add = TRUE, col = "red")
```


### Crop and Mask

```{r, eval=FALSE, include=TRUE}
# crop
crop_dem <- crop(dem_2022, extent(sj))
plot(crop_dem)

# mask
mask_dem <- mask(crop_dem, sj)
plot(mask_dem)

# save 
raster::writeRaster(mask_dem, filename = file.path("Data/DEM or DSM/sj_dem.tif"), format="GTiff", overwrite=TRUE)
```

### Load clipped dem

```{r, eval=FALSE, include=TRUE}
sj_dem <- raster("Data/DEM or DSM/sj_dem.tif")
plot(sj_dem)
```


### Slope

```{r, eval=FALSE, include=TRUE}
slope <- terrain(sj_dem, "slope", unit = "degrees")
plot(slope)

# save 
#raster::writeRaster(slope, filename = file.path("Data/DEM or DSM/sj_slope.tif"), format="GTiff", overwrite=TRUE)
```



### LS from ArcGIS Pro

```{r, eval=FALSE, include=TRUE}
# flow accumulation
ls <- raster("Data/DEM or DSM/LS_1_1/LS_1_1.tif")

plot(ls)
plot(st_geometry(sj), add = TRUE)

# crop
crop_ls <- crop(ls, extent(sj))
mask_ls <- mask(crop_ls, sj)
plot(mask_ls)

mask_ls
ef

# resampling/interpolation
target <- slope
res(target) <- res(ef) # 30 meters
mask_ls <- raster::resample(mask_ls, target, method = "bilinear")
plot(mask_ls)

# save 
raster::writeRaster(mask_ls, filename = file.path("Data/LS /LS_factor.tif"), format="GTiff", overwrite=TRUE)
```


```{r}
# load LS factor
ls_factor <- raster("Data/LS Factor/LS_factor.tif")
plot(ls_factor)
```


## Cover management

### Load data


```{r}
cs <- raster("Data/Crop cover/Reclassify_SJ_Cropcover_2021.tif")
plot(cs)

# crop and mask
crop_cs <- crop(cs, extent(sj))
mask_cs <- mask(crop_cs, sj)
plot(mask_cs)

# crop factor
c_factor <- mask_cs/100
plot(c_factor)
```


## Support Practices: Cropping (P-factor)

- Cmanagementpractices = 1 for conventional Strip cropping;
- Cmanagementpractices = 0.35 for conservation contour cropping; 
- Cmanagementpractices = 0.25 for terrace cropping.


# Erosion calculation

## setup

```{r}
e_factor <- raster::resample(ef, r_moore, method = "bilinear")
ls_factor <- raster::resample(ls_factor, r_moore, method = "bilinear")
c_factor <- raster::resample(c_factor, r_moore, method = "bilinear")
```


## First scenario

### Conventional Strip Cropping

```{r}
RUSLE_1 <- r_moore * e_factor * ls_factor * c_factor * 1
plot(RUSLE_1)

# save 
raster::writeRaster(RUSLE_1, filename = file.path("Data/RUSLE/RUSLE_1.tif"), format="GTiff", overwrite=TRUE)
```


## Second scenario

### Conservation Contour Cropping


```{r}
RUSLE_2 <- r_moore * e_factor * ls_factor * c_factor * 0.35
plot(RUSLE_2)

# save 
raster::writeRaster(RUSLE_2, filename = file.path("Data/RUSLE/RUSLE_2.tif"), format="GTiff", overwrite=TRUE)
```


## Third scenario

### Terrace Cropping


```{r}
RUSLE_3 <- r_moore * e_factor * ls_factor * c_factor * 0.25
plot(RUSLE_3)

# save 
raster::writeRaster(RUSLE_3, filename = file.path("Data/RUSLE/RUSLE_3.tif"), format="GTiff", overwrite=TRUE)
```


## Comparison of scenarios

### RUSLE 1 vs RUSLE 2

```{r}
r1_r2 <- RUSLE_1 - RUSLE_2
plot(r1_r2)
```



# Projections

```{r}
library(wesanderson)
# Gradient color
pal <- wes_palette("Zissou1", 100, type = "continuous")
```


## Load ppt data

```{r}
fut_ppt <- stack("Data/Rainfall/dataset-sis-biodiversity-cmip5-global-83e8ca0e-4eea-4051-a911-c3299d0f2380 (1)/BIO12_gfdl-esm2m_rcp85_r1i1p1_1950-2100_v1.0.nc")
plot(fut_ppt, col = pal)
```


## Referencing

```{r}
library(tigris)
us <- states(cb = TRUE) %>% 
  filter(!NAME %in% c("Puerto Rico", "Guam", "American Samoa", "United States Virgin Islands", "Commonwealth of the Northern Mariana Islands", "Alaska", "Hawaii")) 

plot(fut_ppt[[1]])
plot(st_geometry(us), add = TRUE)
```


## Select study years

- 2030

- 2050

- 2070
  
- 2100


```{r}
# 2030s
y2030 <- fut_ppt[[81]]*3600*24*365*1000
plot(y2030)

# 2050s
y2050 <- fut_ppt[[101]]*3600*24*365*1000
plot(y2050)

# 2070s
y2070 <- fut_ppt[[121]]*3600*24*365*1000
plot(y2070)

# 2100s
y2100 <- fut_ppt[[151]]*3600*24*365*1000
plot(y2100)

# stack
fut_ppt <- stack(y2030, y2050, y2070, y2100)
names(fut_ppt) <- c("2030", "2050", "2070", "2100")
plot(fut_ppt)
```



## Cropping and Masking

```{r, eval=FALSE, include=TRUE}
# cropping
crop_ppt <- raster::crop(fut_ppt, extent(california))
plot(crop_ppt)

# masking
mask_ppt <- raster::mask(crop_ppt, mask = california)
plot(mask_ppt)

# load target layer
res_rnf <- raster("Data/Rainfall/PRISM/res_rnf_california.tif")

# resampling/interpolation
res_ppt <- raster::resample(mask_ppt, res_rnf, method = "bilinear")
plot(res_ppt)

raster::writeRaster(res_ppt, filename = file.path("Data/FUTURE_PROJECTIONS/res_ppt_california.tif"), format="GTiff", overwrite=TRUE)
```


```{r, eval=FALSE, include=TRUE}
res_ppt <- brick("Data/FUTURE_PROJECTIONS/res_ppt_california.tif")
plot(res_ppt)

plot(res_ppt[[1]])
plot(st_geometry(sj), add = TRUE)

# cropping
crop_ppt <- raster::crop(res_ppt, extent(sj))
plot(crop_ppt)

# masking
mask_ppt <- raster::mask(crop_ppt, mask = sj)
plot(mask_ppt)

raster::writeRaster(mask_ppt, filename = file.path("Data/FUTURE_PROJECTIONS/res_ppt_sj.tif"), format="GTiff", overwrite=TRUE)
```


```{r}
res_ppt <- stack("Data/FUTURE_PROJECTIONS/res_ppt_sj.tif")
#res_ppt <- res_ppt*3600*24*365*1000
names(res_ppt) <- c("2030", "2050", "2070", "2100")
plot(res_ppt)

# comparison of decades
ppt_df <- res_ppt %>% 
  as.data.frame(xy = TRUE) %>% 
  pivot_longer(names_to = "key", values_to = "value", cols = c(3:4)) %>% 
  filter(!is.na(value))

ppt_df %>% 
  ggplot() +
  geom_raster(aes(x = x, y = y, fill = value)) +
  scale_fill_distiller(palette = "RdPu", direction = 1) +
  theme_minimal() +
  facet_wrap(~ key)
  

# erosivity factor
r_factor_future <- calc(x = res_ppt, fun = calculate_r_moore)

# setup
r_factor_future <- raster::resample(r_factor_future, r_moore, method = "bilinear")
plot(r_factor_future)
```



## First scenario

### Conventional Strip Cropping

```{r}
RUSLE_1_future <- r_factor_future * e_factor * ls_factor * c_factor * 1
names(RUSLE_1_future) <- c("2030", "2050", "2070", "2100")
plot(RUSLE_1_future)

# save 
raster::writeRaster(RUSLE_1_future, filename = file.path("RUSLE_1_future.tif"), format="GTiff", bylayer = TRUE, overwrite=TRUE)
```


## Second scenario

### Conservation Contour Cropping


```{r}
RUSLE_2_future <- r_factor_future * e_factor * ls_factor * c_factor * 0.35
names(RUSLE_2_future) <- c("2030", "2050", "2070", "2100")
plot(RUSLE_2_future)

# save 
raster::writeRaster(RUSLE_2_future, filename = file.path("RUSLE_2_future.tif"), format="GTiff", bylayer = TRUE, overwrite=TRUE)
```


## Third scenario

### Terrace Cropping


```{r}
RUSLE_3_future <- r_factor_future * e_factor * ls_factor * c_factor * 0.25
names(RUSLE_3_future) <- c("2030", "2050", "2070", "2100")
plot(RUSLE_3_future)

# save 
raster::writeRaster(RUSLE_3_future, filename = file.path("RUSLE_3_future.tif"), format="GTiff", bylayer = TRUE, overwrite=TRUE)
```



# EDA

```{r}
rnf_sj <- raster::resample(rnf_sj, res_ppt, method = "bilinear")
names(rnf_sj) <- "2021"

plot(stack(rnf_sj, res_ppt), col = rev(pal))
```

